Benchmark!!1

### 2021-02-05 paper/main.tex
more paper progress

\section{The benchmarks}
We consider five different benchmark algorithms that perform tasks
commonly encountered in transport optimization. Together they capture
key aspects of computationally intensive tasks in heuristics and
exact methods for transport optimization. Additionally, we believe
that they also capture such aspects for other fields of application
of operations research, such as e.g. scheduling. For the sake of
simplicity, we consider the symmetric travelling salesperson problem
(TSP) as a base problem, i.e. input data are in the form of a
symmetric distance matrix while a solution is a permutation. However
some benchmark algorithms solve different problems than the TSP based on these
TSP data, as described below.

@@

\subsection{ESPPRC: Dynamic programming for column generation}
Many state-of-the-art exact methods for vehicle routing problems rely
on column generation (see e.g.~\cite{Baldacci:2012survey}). In such
methods, a significant amount of CPU effort is spent solving the
\emph{pricing subproblem}, which differs depending on the problem at
hand but is nonetheless often an elementary shortest path problem with
resource constraints (ESPPRC). While it can be solved heuristically,
in order to establish optimality it eventually needs to be solved
exactly as well. For this benchmark we implement a dynamic
programming algorithm similar to the one by~\cite{feillet:2004exact}
to solve the ESPPRC, albeit without time windows. One difficulty in
this problem stems from the existence of negative-cost cycles in the
graph.

In order to simulate an environment similar to the one encountered in
column generation for VRPs, we derive an ESPPRC instance from each TSP
solution by using an auxilliary distance matrix $c'$ from the
existing distance matrix $c$, as follows:
\begin{enumerate}
\item For each vertex $j$, let $\pi_j$ be the cost of arc $(i, j)$,
  where $i$ is the predecessor of $j$ in the TSP solution.
\item For each arc $(i, j)$, set $c'_{ij}$ to $c_{ij} - d_j$.
\end{enumerate}
Additionally, we generate resource consumption as follows: we consider
6 resources numbered from 0 to 5, and any vertex $i$ consumes resource
$r$ iff the $r^{th}$ bit of $i$'s binary representation is 1
(considering that the first bit is bit 0). That is, in C syntax, iff
$i & (1 << r) > 0$. Capacity is set to 1 for each resource.
The algorithm then computes the elementary shortest path with resource
constraints on $c'$ from 0 to 0 (1 to 1 in Julia).

@@

Each label needs (i) a reference to its predecessor, so that
paths can be reconstructed and (ii) a collection of references to
its successors, in order to recursively propagate dominance and later
ignore labels that are successors of an already proved-to-be-dominated
label. Rust does not allow that, at least not in the intuitive way
where a label is encapsulated in a \verb|struct|, including references
to predecessor and successors. This is because Rust is focused on
security, while this type of construct can lead to unsafe operations if
not used properly. This means that this benchmark cannot be
implemented with Rust, even though it is consistent and reliable with all
other implementations. In order to remediate this issue, we also
implement a second benchmark to implement the same algorithm, although
in a less intuitive way: all labels are stored in a collection, and
the indices in this collection are used in other collections to store all
predecessors and successors. This second implementation is more
tedious, but it may also be more efficient, so we also implement it in
other languages. We call it \emph{espprc-index}.

@@

\subsection{Maxflow: Maximum flow problem}
Another very popular framework for the exact solution of routing
problems is branch and cut. \cite{padberg:532}

Similar to the ESPPRC benchmark above and for the same reason, we use
floating-point number representation for this benchmark.

The checksum for a given permutation seed is the maximum flow value
obtained when using that permutation seed to generate the MFP
instance, truncated to its integer part.

>>>>

\section{The benchmarks}
We consider five different benchmark algorithms that perform tasks
commonly encountered in transport optimization. Together they capture
key aspects of computationally intensive tasks in heuristics and
exact methods for transport optimization, although they do not
necessarily use the most efficient algorithm to solve the problem they
are tackling. Additionally, we believe
that they also capture such aspects for other fields of application
of operations research, such as e.g. scheduling. For the sake of
simplicity, we consider the symmetric travelling salesperson problem
(TSP) as a base problem, i.e. input data are in the form of a
symmetric distance matrix while a solution is a permutation. However
some benchmark algorithms solve different problems than the TSP based on these
TSP data, as described below.

@@

\subsection{Dynamic programming for column generation}
Many state-of-the-art exact methods for vehicle routing problems rely
on column generation (see e.g.~\cite{Baldacci:2012survey}). In such
methods, a significant amount of CPU effort is spent solving the
\emph{pricing subproblem}, which differs depending on the problem at
hand but is nonetheless often an elementary shortest path problem with
resource constraints (ESPPRC). While it can be solved heuristically,
in order to establish optimality it eventually needs to be solved
exactly as well. For this benchmark we implement a dynamic
programming algorithm similar to the one by~\cite{feillet:2004exact}
to solve the ESPPRC, albeit without time windows. One difficulty in
this problem stems from the existence of negative-cost cycles in the
graph.

In order to simulate an environment similar to the one encountered in
column generation for VRPs, we derive an ESPPRC instance from each TSP
solution by using an auxilliary distance matrix $c'$ from the
distance matrix $c$, as follows:
\begin{enumerate}
\item For each vertex $j$, let $\pi_j$ be the cost of arc $(i, j)$,
  where $i$ is the predecessor of $j$ in the TSP solution.
\item For each arc $(i, j)$, set $c'_{ij}$ to $c_{ij} - d_j$.
\end{enumerate}
Additionally, we generate resource consumption as follows: we consider
6 resources numbered from 0 to 5, and any vertex $i$ consumes resource
$r$ iff the $r^{th}$ bit of $i$'s binary representation is 1
(considering that the first bit is bit 0). That is, in C syntax, iff
$i & (1 << r) > 0$. Capacity is set to 1 for each resource.
The algorithm then computes the elementary shortest path with resource
constraints on $c'$ from 0 to 0 (1 to 1 in Julia).

@@

Each label needs (i) a reference to its predecessor, so that
paths can be reconstructed and (ii) a collection of references to
its successors, in order to recursively propagate dominance and later
ignore labels that are successors of an already proved-to-be-dominated
label. Rust does not allow that, at least not in the intuitive way
where a label is encapsulated in a \verb|struct|, including references
to predecessor and successors. This is because Rust is focused on
security, while this type of construct can lead to unsafe operations if
not used properly. This means that this benchmark cannot be
implemented with Rust, even though it is consistent and reliable with all
other implementations. In order to remediate this issue, we also
implement a second benchmark to implement the same algorithm, although
in a less intuitive way: all labels are stored in a collection, and
the indices in this collection are used in other collections to store all
predecessors and successors. This second implementation represents a
different benchmark, which we call \emph{espprc-index}, whereas the
base benchmark is called \emph{espprc}. The version with indices is more
tedious to implement, but it may also be more efficient, so we also
implement it in other languages.

@@

\subsection{Maximum flow problem}
Another popular framework for the exact solution of routing
problems is \emph{branch and cut}. Branch and cut is typically applied
to solve mixed-integer linear programs; it is akin to branch and bound
but only considers a subset of all constraints explicitly. Other
constraints are generated dynamically when they are found to be
violated while exploring the search tree. These constraints are said
to be \emph{separated}. One early success story of branch and cut is
actually in application to the symmetric TSP~\citep{padberg:532}, by
separating subtour elimination constraints. The separation procedure
involves finding a minimum capacity cut, which itself involves solving
a maximum flow problem. For this reason the maximum flow problem is
important in vehicle routing, and we dedicate one benchmark to solving
it. The \emph{maxflow} benchmark implements the Edmonds-Karp
algorithm~\cite{edmonds-karp}, which is a specific version of the
Ford-Fulkerson~\cite{ford-fulkerson} augmenting path method. The
specificity resides in how the augmenting paths are generated.

We derive a capacity graph $C$ from the distance matrix $c$ and a
given TSP solution as follows: 
\begin{enumerate}
\item For each vertex $j$, let $t_j$ be the cost of arc $(i, j)$,
  where $i$ is the predecessor of $j$ in the TSP solution.
\item For each arc $(i, j)$, set $C_{ij}$ to $c_{ij}/1000$ if $c_{ij}
  > t_j$, otherwise set it to 0.
\end{enumerate}

For any starting seed, the capacity graph is generated then the
Edmonds-Karp algorithm is applied with 0 as \emph{source} and every other
node as \emph{sink}. The checksum for that permutation seed is the sum
of all maximum flow values thus obtained, truncated to its integer part.

The reason why capacity values are distance values divided by 1000 is
that otherwise, checksum values can exceed the largest possible
integer representation with 32 bits, which can be a problem with some
languages. This means that floating-point numbers are used, which is
representative of the use case of branch and cut for vehicle routing problems.
